---
title: "Regmods-016 Assignment"
author: "Lili Sun"
date: "21/12/2014"
output:
  pdf_document:
    fig_caption: yes
    fig_width: 4
    keep_tex: no
documentclass: article
classoption: a4paper
---

# Executive Summary

When considering mpg and am alone, we fit it to a linear model and find the coefficient is about 7.245. That shows the manual transmission cars are 7.2 mpg more efficient than the automatics.
When considering mpg with other variables, we found the most significant varible is `am`


# Basic Analysis

```{r "t.test", echo=FALSE}
data(mtcars)
am_factor <- as.factor(mtcars$am)
levels(am_factor) <- c("Automatic","Manual")
t1 <- t.test(mtcars$mpg ~ am_factor)
```
The relationship of interest is that between `mpg` and transmission `am`. The boxplot (appendix 1, fig. 1) shows a clear increase in the median mpg for manual transmission over automatic transmission. The normal q-q plots (appendix 1, fig. 2) show that the mpg of automatic and manual transmission cars in the data set is approximately normally distributed. The Welch Two Sample t-test suggests a significant difference between the means of the two groups (P = `r sprintf("%0.4f", t1$p.value)`, two sided).

# Simple Model

```{r}
simple <- lm(mpg ~ am, data = mtcars)
simple$coefficients
```
The relationsip shown above can be expressed as a simple linear model: $mpg = 7.245am + 17.147$. The model resulting in an intercept of 17.147, which is the mean mpg of automatic type. The coefficient is 7.245, which is the increase of manual type over automatic type. 

# Full Model

```{r}
full <- lm(mpg ~ ., data = mtcars)
full$coefficients
```
The relationship shown above can be expressed as simple linear model:
$mpg = \Sigma_i\beta_ix_i + c$, here $x_i$ is the variables, and $\beta_i$ is their responding coefficient.
As shown above, the most significant variable is `am`.

# Stepwise Regression
```{r}
stepFarward <- step(simple, direction="forward", trace = 0,  scope=(~ cyl + disp + hp + drat + wt + qsec + vs + am + gear + carb))
stepBackward <- step(full, direction = "backward", trace = 0)
stepFarward$coefficients
stepBackward$coefficients
```
We use forward and backward regressor selection. The forward results in a model: `mpg ~ am+hp+wt+qsec`. While the backward results in a model: `mpg ~ wt+qsec+am`

# Model Selection
Here we use the result of farward reression.

# Diagnostics
```{r}
round(dfbetas(stepFarward),4)
round(hatvalues(stepFarward),4)
```
\pagebreak

# Appendix 1: Plots

## Simple comparison

```{r "boxplot", echo=FALSE, fig.cap="Boxplot for MPG by am"}
boxplot(mtcars$mpg ~ mtcars$am)
```

## Normal q-q plot

```{r "qqplot", echo=FALSE, fig.cap='Normal q-q plot for MPG by transmission type', fig.width=8}
par(mfrow = c(1,2)) 
qqnorm(scale(mtcars$mpg[mtcars$am == "0"]), main = "Automatic")
abline(0,1)
qqnorm(scale(mtcars$mpg[mtcars$am == "1"]), main = "Manual")
abline(0,1)
```




\pagebreak

## Diagnostic plots

```{r "diagnostics", echo=FALSE, fig.width=8, fig.height=8, fig.cap="Model diagnostics plots"}
#par(mfrow = c(2, 2))
#plot(mInt)
```




